---
title: "Vital status INT and YEAR inconsistencies"
author: "Haley Hunter-Zinck"
date: "`r Sys.setenv(TZ='America/Los_Angeles'); format(Sys.time(), '%B %d, %Y at %I:%M %p %Z')`"
output: html_document
parameters:
  synid_file_patient: x
  synid_version_patient: x
---

## Overview

We look for discrepancies in vital status information submitted for main GENIE consortium releases.  

We check the following pairs of variables:

* YEAR_CONTACT and YEAR_DEATH
* YEAR_DEATH and INT_DOD
* YEAR_CONTACT and INT_CONTACT

for inconsistencies with any of the following rules: 

1. If the YEAR_CONTACT > YEAR_DEATH
2. If YEAR_DEATH or INT_DOD are greater than 365 days apart  
3. If YEAR_CONTACT or INT_CONTACT are greater than 365 days apart  

## Parameters

Synapse ID of patient file: `r synid_file_patient`

Synapse version of patient file: `r synid_version_patient`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(synapser)
invisible(synLogin())
library(glue)
library(dplyr)

```

```{r functions, include = F}
#' Download and load data stored in csv or other delimited format on Synapse
#' into an R data frame.
#' 
#' @param synapse_id Synapse ID
#' @version Version of the Synapse entity to download.  NA will load current
#' version
#' @param set Delimiter for file
#' @param na.strings Vector of strings to be read in as NA values
#' @param header TRUE if the file contains a header row; FALSE otherwise.
#' @param check_names TRUE if column names should be modified for compatibility 
#' with R upon reading; FALSE otherwise.
#' @return data frame
get_synapse_entity_data_in_csv <- function(synapse_id, 
                                           version = NA,
                                           sep = ",", 
                                           na.strings = c("NA"), 
                                           header = T,
                                           check_names = F, 
                                           comment.char = "#") {
  
  if (is.na(version)) {
    entity <- synGet(synapse_id)
  } else {
    entity <- synGet(synapse_id, version = version)
  }
  
  data <- read.csv(entity$path, stringsAsFactors = F, 
                   na.strings = na.strings, sep = sep, check.names = check_names,
                   header = header, comment.char = comment.char)
  return(data)
}
```

```{r read, include = F}

pat <- get_synapse_entity_data_in_csv(synid_file_patient, version = synid_version_patient, sep = "\t")

```


## Vital data summary

Number of patients = `r nrow(pat)` 

## Checks for vital INT and YEAR inconsistencies

### Check 1

```{r contact after death, include = F}
df_year_invalid <- pat %>% 
  filter(grepl(pattern = "^[0-9]{4}$", x = YEAR_CONTACT) & grepl(pattern = "^[0-9]{4}$", x = YEAR_DEATH) & YEAR_CONTACT > YEAR_DEATH) %>%
  select(PATIENT_ID, CENTER, YEAR_CONTACT, YEAR_DEATH)
```

Number of cases with contact after death: `r nrow(df_year_invalid)`

Number of cases with contact after death by center:
```{r contact after death by center, include = T}
print(df_year_invalid %>% group_by(CENTER) %>% count() %>% arrange(CENTER))
```

### Check 2

```{r death int and year, include = F}
df_int_death <- pat %>%
  mutate(INT_DOD_EST = (as.double(YEAR_DEATH) - as.double(BIRTH_YEAR)) * 365) %>%
  mutate(INT_DOD_DIFF = abs(as.double(INT_DOD) - INT_DOD_EST)) %>%
  filter(INT_DOD_DIFF > 365) %>%
  select(PATIENT_ID, CENTER, YEAR_DEATH, BIRTH_YEAR, INT_DOD, INT_DOD_EST, INT_DOD_DIFF)
```

Number of cases with inconsistent INT_DOD and YEAR_DEATH: `r nrow(df_int_death)`

Number of cases with inconsistent INT_DOD and YEAR_DEATH by center:
```{r death int and year by center, include = T}
print(df_int_death %>% group_by(CENTER) %>% count() %>% arrange(CENTER))
```

### Check 3

```{r contact int and year, include = F}
df_int_contact <- pat %>%
  mutate(INT_CONTACT_EST = (as.double(YEAR_CONTACT) - as.double(BIRTH_YEAR)) * 365) %>%
  mutate(INT_CONTACT_DIFF = abs(as.double(INT_CONTACT) - INT_CONTACT_EST)) %>%
  filter(INT_CONTACT_DIFF > 365) %>%
  select(PATIENT_ID, CENTER, YEAR_CONTACT, BIRTH_YEAR, INT_CONTACT, INT_CONTACT_EST, INT_CONTACT_DIFF)
```

Number of cases with inconsistent INT_CONTACT and YEAR_CONTACT: `r nrow(df_int_contact)`

Number of cases with inconsistent INT_CONTACT and YEAR_CONTACT by center:
```{r contact int and year by center, include = T}
print(df_int_contact %>% group_by(CENTER) %>% count() %>% arrange(CENTER))
```
