---
title: "BPC NSCLC - immunotherapy use"
subtitle: "Using '3.1 consortium' (Phase 1 and Phase 2 combined)"
author: "Alex Paynter"
date: "`r Sys.Date()`"
output:
  quarto::quarto_html:
    toc: true
    toc_float: true
    theme:
      name: paper
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(purrr); library(here); library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source) # also load
```


```{r}
dft_ca_ind <- readr::read_csv(here('data-raw', "cancer_level_dataset_index.csv"),
                           show_col_types = F)

dft_ca_non_ind <- readr::read_csv(here('data-raw', "cancer_level_dataset_non_index.csv"),
                           show_col_types = F)

dft_reg <- readr::read_csv(here('data-raw', "regimen_cancer_level_dataset.csv"),
                           show_col_types = F)

dft_drug <- dft_reg %>%
  create_drug_dat(.)

dft_phase <- rbind(
  select(dft_ca_ind, record_id, ca_seq, phase),
  select(dft_ca_non_ind, record_id, ca_seq, phase)
)

dft_drug %<>%
  left_join(
    .,
    dft_phase,
    by = c('record_id', 'ca_seq')
  )

dft_drug_ind <- dft_ca_ind %>%
  select(record_id, ca_seq) %>%
  left_join(., dft_drug, by = c("record_id", "ca_seq"))




vec_io <- c(
  "Atezolizumab",
  "Durvalumab",
  # Don't see anything resembling Cemiplimab-rwlc
  "Nivolumab",
  "Pembrolizumab",
  "Ipilimumab",
  "Tremelimumab"
)

sum_help <- function(dat) {
  dat %>%
    summarize(
      n_people = length(unique(record_id)),
      n_cases = length(unique(paste0(record_id, "Q", ca_seq))),
      .groups = "drop"
    )
}

ft_help <- function(dat) {
  dat %>%
    flextable(.) %>%
    autofit(.)
}

```

## Query

From Tony Wu:  Can you perform a count for the BPC NSCLC Phase 2 cohort for all patients that received an I/O drug?

The list of IO drugs provided:

- Anti-PD-1
    - Atezolizumab (Tecentriq)
    - Durvalumab (Imfinzi)
    - Cemiplimab-rwlc (Libtayo)
    - Nivolumab (Opdivo)
    - Pembrolizumab (Keytruda)

- Anti-CTLA4
    - Ipilimumab (Yervoy)
    - Tremelimumab (Imjudo)
    
## Analysis 

The dataset analyzed is called "3.1 consortium" which I'm told is Phase 1 and 2 combined for the NSCLC cohort.  `n_people` is the number of unique record IDs with a regimen containing a drug on the list. `n_cases` is the same for the number of unique {record_id, ca_seq} pairs (i.e. someone could have a primary in the left and right lung both counted).

The first set of results I state filters to index cancer cases, which I would recommend.  The results with non-index cancer cases almost certainly double-count regimens for the `n_cases` numbers because regimens are not always uniquely assigned to a cancer.  But they're there if that's what you wanted, and might be valid for `n_people`.

I did not see any agents resembling "Cemiplimab-rwlc" or "Tremelimumab" in the data, so tell me if that sounds off.

## Results

Overall results:

```{r, include = T}
dft_drug_ind %>%
  filter(agent %in% vec_io) %>%
  sum_help(.) %>%
  ft_help(.)
```

Overall results split by phase:

```{r, include = T}
dft_drug_ind %>%
  filter(agent %in% vec_io) %>%
  group_by(phase) %>%
  sum_help(.) %>%
  ft_help(.)
```

Overall results split by drug ('agent'):

```{r, include = T}
dft_drug_ind %>%
  filter(agent %in% vec_io) %>%
  group_by(agent) %>%
  sum_help(.) %>%
  ft_help(.)
```

## Results including non-index cancers

Overall results including non-index cancers:

```{r, include = T}
dft_drug %>%
  filter(agent %in% vec_io) %>%
  sum_help(.) %>%
  ft_help(.)
```

Overall results split by phase including non-index cancers:

```{r, include = T}
dft_drug %>%
  filter(agent %in% vec_io) %>%
  group_by(phase) %>%
  sum_help(.) %>%
  ft_help(.)
```

Overall results split by drug ('agent') including non-index cancers:

```{r, include = T}
dft_drug %>%
  filter(agent %in% vec_io) %>%
  group_by(agent) %>%
  sum_help(.) %>%
  ft_help(.)
```