---
title: "wu-brca-main-genie-request"
author: "Alex Paynter"
date: "Updated July 25, 2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F)
```

```{r}
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(purrr)
library(magrittr)
library(glue)
library(fs)
library(here)
library(janitor) # interactive work

library(ggplot2)
library(plotly)
library(knitr)
library(khroma) # paul tol palettes.
```


```{r}
purrr::walk(fs::dir_ls(here('R')), .f = source)
```

```{r}
dft_releases <- readr::read_rds(file = here('data', 'releases_by_pt_nested.rds'))
```

```{r}
count_patients <- function(dat) {
  rtn <- dat %>%
    group_by(center, oncotree_primary_node) %>%
    summarize(
      n_total = n(),
      n_variant = sum(variant),
      .groups = "drop"
    )
  
  return(rtn)
}

# Example of this running in one row:
# test_df <- dft_releases %>% slice(1) %>% pull(release_sum) %>% `[[`(1)
# count_patients(test_df)

dft_mut_counts <- dft_releases %>%
  mutate(
    sum = purrr::map(.x = release_sum, .f = count_patients)
  ) %>%
  select(-release_sum) %>%
  unnest(sum)

# Request was to isolate the 4 "known" impact cancers:
# Breast (BREAST)
# Ovarian (OVARY and OVARY/FALLOPIAN_TUBE)
# Prostate (PROSTATE)
# Pancreatic (PANCREAS)
vec_known_brca <- c("BREAST", "OVARY", "OVARY/FALLOPIAN_TUBE", "PROSTATE",
                    "PANCREAS")
dft_mut_counts %<>%
  mutate(known_cancers = if_else(oncotree_primary_node %in% vec_known_brca,
                                 T, F, F)) 

dft_mut_counts %<>%
  rename(release = minor) %>%
  mutate(
    release_date = str_sub(minor_createdOn, 1, 10),
    release_date = lubridate::ymd(release_date),
    tissue = str_replace_all(oncotree_primary_node, "_", " "),
    tissue = str_to_title(tissue)
  ) %>%
  select(release, release_date, center, tissue, n_total, n_variant, known_cancers)

vec_cancer_types <- c(
  "Breast", "Ovarian", "Prostate", "Pancreatic", "Other"
)

dft_mut_counts <- dft_mut_counts %>%
  mutate(
    cancer_groups = case_when(
      is.na(tissue) ~ vec_cancer_types[5],
      tissue %in% c("Ovary", "Ovary/Fallopian Tube") ~ vec_cancer_types[2],
      tissue %in% "Breast" ~ vec_cancer_types[1],
      tissue %in% "Prostate" ~ vec_cancer_types[3],
      tissue %in% "Pancreas" ~ vec_cancer_types[4],
      T ~ vec_cancer_types[5]
    ),
    cancer_groups = factor(cancer_groups, levels = vec_cancer_types)
  )

# We decided to look only at the consortium releases.  Save the original just in case.
dft_mut_counts_incl_public <- dft_mut_counts
dft_mut_counts <- dft_mut_counts %>%
  filter(!str_detect(release, "-public$"))
```


```{r}
pal <- color("bright")(7)

dft_totals <- dft_mut_counts %>%
  # Implicitly sums over center and tissue type rows:
  group_by(release, release_date) %>%
  summarize(
    n_brca = sum(n_variant), 
    n_total = sum(n_total),
    .groups = "drop"
  )

dft_totals %<>%
  mutate(
    plotly_lab = glue(
      "Release Title: {release}
  Release Date: {release_date}
  Total Patients: {n_total}
  Patients with BRCA mutation: {n_brca}"
    )
  )

gg_total_brca <- ggplot(
  data = dft_totals,
  aes(x = release_date, y = n_brca, text = plotly_lab, group = 1)
  ) + 
  geom_line(color = pal[1], linewidth = 0.75) +
  theme_classic() +
  labs(y = "Patients", x = "Release Date") + 
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

plotly_total_brca <- ggplotly(gg_total_brca, tooltip = "text")

gg_total_all <- ggplot(
  data = dft_totals,
  aes(x = release_date, y = n_brca, text = plotly_lab, group = 1)
  ) + 
  geom_line(color = pal[2], linewidth = 0.75) +
  theme_classic() +
  labs(y = "Patients", x = "Release Date") + 
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

plotly_total_all <- ggplotly(gg_total_all, tooltip = "text")
```





```{r}
dft_mut_counts_by_group <- dft_mut_counts %>% 
  group_by(release, release_date, cancer_groups) %>%
  summarize(n_brca = sum(n_variant), .groups = "drop")

dft_mut_counts_by_group %<>%
  # apparent data errors:
  filter(!(release %in% c("13.2-consortium", "8.6-consortium", "14.0-public")))
```

```{r}
library(khroma)
colour("bright")
gg <- ggplot(dft_mut_counts_by_group,
       aes(x = release_date, y = n_brca, color = cancer_groups)) + 
  geom_line() + 
  theme_bw() + 
  scale_color_vibrant(name = NULL) + 
  labs(y = "Patients", x = "Release Date") + 
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

gg
# dft_mut_counts_by_group %>%
#   filter(cancer_groups %in% "Other") %>% 
#   arrange(release_date) %>%
#   print(n = 500)

other_cancers <- dft_mut_counts %>% 
  filter(cancer_groups %in% "Other") %>%
  pull(tissue) %>%
  unique %>%
  sort %>%
  paste(collapse = ", ")

```



## Introduction

This report examines those with BRCA1 or BRCA2 mutations in the main GENIE database.  Several important notes:

1. Only somatic mutations are included.
2. Cancers are classified according to the oncotree primary node (Breast, Pancreatic, etc).  For any without a designation we coded them as "Other".
3. "Release date" is currently when the object was created in synapse.  This isn't ideal - work in progress.
4. Two releases were excluded because they seem erroneous.
5. Line jitter:  Private and public releases are included, and retractions happen over time, so the lines jitter as they go a bit.

## Accruals over time

Other cancers here includes: `r other_cancers`

```{r, include = T}
gg
```

Numbers from latest release:

```{r, include = T}
dft_mut_counts_by_group %>% 
  filter(release %in% "14.6-consortium") %>%
  kable(.)
```


```{r}
dft_pros_chk <- dft_mut_counts_by_group %>%
  filter(cancer_groups %in% "Prostate") %>%
  select(release, release_date, n_brca) %>%
  arrange(release_date)
```

# Prostate numbers by release

The following table is provided for possible validation with previous work (Jen Hoppe may have done something similar for the prostate group).

`n_brca` is the number of participants with a BRCA1 or BRCA2 somatic mutation.

```{r, include = T}
dft_pros_chk %>% huxtable::huxtable(.)
```

