---
title: "Tony Wu Prostate BPC request - May 9"
author: "Alex Paynter"
date: "5/9/2023"
output: 
  html_document:
    toc: false
    theme: paper
editor_options: 
  chunk_output_type: console
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      warning = F, message = F,
                      fig.width = 7, fig.height = 5)
k_dpi <- 150
```


<style type="text/css">
.main-container {
  max-width: 1000px !important;
  margin: auto;
}
</style>



```{r}
library(cli)
library(readr)
library(here)
library(dplyr)
library(tidyr)
library(purrr)
library(rlang)
library(magrittr)
library(janitor)
library(stringr)
library(vctrs)
library(glue)
library(lubridate)
library(huxtable)


# Source everything in the "R" folder - just functions, no scripts.
purrr::walk(.x = here("R", dir(here("R"))), .f = source)
```


```{r, warning = T}
read_wrap <- function(p) {
  read_csv(file = here("data-raw", p), show_col_types = F)
}

dft_pt <- read_wrap("patient_level_dataset.csv")
dft_ca_ind <- read_wrap("cancer_level_dataset_index.csv")
dft_med_onc <- read_wrap("med_onc_note_level_dataset.csv")

# A few sanity checks on the data:
if ((dft_pt$record_id %>% duplicated %>% any)) {
  stop("Duplicated records in patient level dataset.")
}

if (any(dft_pt$naaccr_sex_code != "Male")) {
  warning("Non-male sex detected for at least one participant (unexpectedly for Prostate cancer)")
}

```


## Question

From Tony: "looking to capture number of patients who were designated castrate resistant after first metastasis from the total cohort and the numbers from just 2018 alone"

Clarifications:

- This is designated castrate resistant **for the first time**.
- Fine to use guesses on birthdays for the 2018 estimate.
- This analysis includes both those who were diagnosed at stage IV with dmets, and those with a distant metastasis noted during followup.

```{r}
# Functions borrowed from landscape manuscript which get rid of repeats,
#   handle the different stages, etc.

dft_cast_block <- make_cast_status_block(
  dft_med_onc, 
  dft_ca_ind, 
  remove_hspc_after_crpc = T
) %>%
  filter(md_cast_status_f %in% "Castrate-Resistant") %>%
  select(record_id, crpc_yrs = dx_block_start)

dft_dmet <- make_dmet_status_block(
  dft_ca_ind
) %>%
  filter(dmet_status %in% "Distant Metastasis") %>%
  select(record_id, dmet_yrs = dx_block_start)


dft_req <- left_join(
  dft_dmet,
  dft_cast_block,
  by = "record_id"
) %>%
  # add in the data we need for birthdays.
  left_join(
    .,
    select(dft_pt, record_id, birth_year),
    by = "record_id"
  ) %>%
  left_join(
    .,
    select(dft_ca_ind, record_id, dob_ca_dx_yrs),
    by = "record_id"
  )


dft_req %<>%
  filter(crpc_yrs >= dmet_yrs) %>%
# base_crpc_date is the diagnosis date assuming everyone was born on
#   January 1st.
  mutate(
    base_crpc_date = ymd(paste0(birth_year,"-01-01")) +
      dyears(dob_ca_dx_yrs) +
      dyears(crpc_yrs)
    ) 
```


```{r}
n_total_crpc_after_dmet <- dft_req %>%
  nrow(.)
```

## Overall cohort

The number of people with their first CRPC designation on or after the first distant metastasis is `r n_total_crpc_after_dmet`.

```{r, do_boostrap}
one_iter_yr <- function(s, dat) {
  set.seed(s)
  dat %>%
    mutate(
      add_days = sample.int(365, n(), replace = T) - 1,
      crpc_date = base_crpc_date + ddays(add_days),
      crpc_yr = year(crpc_date)
    ) %>%
    count(crpc_yr)
}

n_rep <- 10^4 # takes about 60s to run this way.
reps <- tibble(id = 1:n_rep) %>%
  mutate(s = sample.int(10^6, size = n(), replace = T)) %>%
  mutate(dat = purrr::map(.x = s,
                          .f = one_iter_yr,
                          dat = dft_req)) %>%
  unnest(dat) %>%
  mutate(crpc_yr = factor(crpc_yr)) %>%
  tidyr::complete(id, crpc_yr, fill = list(n = 0)) 

rep_sum <- reps %>%
  group_by(crpc_yr) %>%
  summarize(
    mean_n = mean(n, na.rm = T),
    lcb_n = quantile(n, probs = c(0.025)),
    ucb_n = quantile(n, probs = c(0.975)),
    yr_ct_sanity = n(), # does not not need to be output.
  ) %>%
  mutate(n = cfmisc::est_int_str(est = mean_n, 
                                 lower = lcb_n, 
                                 upper = ucb_n,
                                 est_digits = 1,
                                 plus_prefix = F))

# check the yr_ct_sanity if you want here.
rep_sum %<>%
  select(crpc_yr, n)

```

## By Year

We only have birth year, not birth day, for our participants.  One approach to deal with this is to randomly assign everyone a birthday, then bootstrap to get an estimate (repeat this thousands of times and take a mean).  That's what we'll show below.

The estimate for 2018 is `r rep_sum %>% filter(crpc_yr %in% 2018) %>% pull(n)`, where the parenthesis are a 95% confidence interval depending on where the birthdays of our participants fall.  

The table for every year follows:


```{r,include = T}
rep_sum %>%
  huxtable::huxtable(.) %>%
  theme_compact(.)
```

